This is an example of article title

### 2025-11-13 paper/paper.tex
minor improvements

\begin{abstract}
    \href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}} is a
    Julia package that provides an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions. The package also offers a well-documented, performant, and well-tested library of existing Lie groups, their algebra and corresponding group actions.

@@

\subsection{Lie groups}
We denote a Lie group by \(\mathcal{G} = (\mathcal{M}, \cdot)\) where \(\mathcal{M}\) is a smooth manifold and \(\cdot\) is the group operation.
A smooth manifold \(\mathcal{M}\) is a topological space that is locally isomorphic to an Euclidean space \(\mathbb{R}^n\) for some \(n \in \mathbb{N}\), but globally may have a different topology.
We call \(n\) the dimension of the manifold \(\mathcal{M}\), denoted by \(\dim(\mathcal{M}) = n\).
As an example, take the \(2\)-dimensional sphere
\begin{equation*}
\mathbb{S}^2 = \bigl\{ p \in \mathbb{R}^3 \bigm| \lVert p\rVert = 1\bigr\},
\end{equation*}
% MB: maybe it would be better to replace it with an example that can have a group structure? For example the 3-dimensional sphere?

which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
Finally we denote the tangent space at a point \(p \in \mathcal{M}\) by \(T_p\mathcal{M}\). This can be thought of as all “velocities” (direction and speed) in which a curve can “pass through” a point. Formally it is set the equivalence classes of derivatives of smooth curves. Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces
\begin{equation*}
    T\mathcal M = \dot\bigcup_{p \in \mathcal M} T_p\mathcal M
\end{equation*}
the \term{tangent bundle} of $\mathcal M$.
\\
A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\). Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is, for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle. % chktex 9ok
The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$). % -(-\pi) would be out of the range

\subsection{The Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} $\frak g$ of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} defined by $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(e)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
.
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.

@@

The interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types. To generate a point in a specific representation of type \verb|T|, one can use \verb|identity_element(G, T)|.

@@

Alternatively to points, also different manifolds can be used. This is for example the case for the circle group \(\mathbb{S}^1\), which can be represented as angles, points on the unit circle in \(\mathbb{R}^2\) or as complex numbers with unit norm, all three of which are different manifolds to be used internally. To access the underlying manifold, the function \verb|base_manifold(G)| is provided.

The following functions are available and pass directly to the underlying manifold interface from \verb|ManifoldsBase.jl|:
\verb|is_approx(G, g, h)| to check for (approximate) equality of two points,
\verb|is_point(G, g; error=:none)| to check if a point is a valid point on the manifold, where the keyword can be used to throw an \verb|:error|, a \verb|:warn| or an \verb|:info|.
Similarly
\verb|manifold_dimension(G)| to get the dimension of the manifold,
\verb|project(G, q)|, %
%\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
\verb|rand(G)| and \verb|rand(rng, G)| %
%\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
to sample a random point from the manifold using a random number generator \verb|rng| also pass directly on to the manifold.
To access the underlying manifold of the Lie group one can use \verb|base_manifold(G)|.

@@

The identity element can be represented as a value of a special type, \verb|e = Identity(G)| or \verb|Identity(op)| where \verb|op| is the group operation of \verb|G|. This allows for dispatching on the type when defining functions and avoiding unnecessary allocations. If the actual value is needed, one can call

\verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group. The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|%
%\footnote{The in-place variant \texttt{identity\_element!(G, e::T)} infers the representation from the type \texttt{T} of \texttt{e}.}
.

The two central functions are \verb|compose(G, g, h)|
%\footnote{in-place variant \texttt{compose!(G, k, g, h)}.}
to compute the group operation \(g \cdot h\) for two points \(g, h \in \mathcal{G}\) and the inverse \verb|inv(G, g)|
%\footnote{in-place variant \texttt{inv!(G, h, g)}.}
which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).

Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|%
%\footnote{in-place variant \texttt{conjugate!(G, k, g, h)}.}
.

@@

Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|
%\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.}
as well as the combinations \verb|inv_left_compose(G, g, h)|
%\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}}
for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|
%\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}
for computing $g \cdot h^{-1}$. All three fall back to the previously defined functions, but provide an interface to possibly implement more efficient variants in case such exist.

@@

Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code%
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}%
.
Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

@@

Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b, T)| specifying the tangent vector type%
%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
.
Given a tangent vector $X \in \mathfrak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\mathfrak g$\verb|, X, B)|%
%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, b)}}
, where the Basis is again optional.

@@

\subsubsection*{Push forward and pull back of tangent vectors}
To the representation of a tangent vector $X \in \mathfrak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$. This is implemented as \verb|push_forward_tangent(G, g, X)|%
%\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
.
Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\mathfrak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$. This is implemented as \verb|pull_back_tangent(G, g, Y)|%
%\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
.

\subsubsection*{Jacobians}
On Euclidean space, the terms differential and Jacobian are often used interchangeably. Within differential geometry, the differential (or push forward) of a smooth map $f\colon \mathcal M \to \mathcal N$ between two manifolds is a map between the corresponding tangent spaces. We denote it by $Df(p)\colon T_p\mathcal M \to T_{f(p)}\mathcal N$, where $p \in \mathcal M$.
The differential is a linear map between the tangent spaces.
However, in the coordinate-free representation of tangent vectors, there is no matrix representation of this linear map.
But as soon as we choose bases for the tangent spaces and we represent tangent vectors in coordinates (cf.~\verb|get_coordinates| previously), we can represent the differential as a matrix, called the Jacobian matrix.

Currently for both \verb|exp| as well as \verb|conjugate|, the Jacobian is implemented. Note that for \verb|exp|, the domain is the Lie algebra. Since that is a vector space, its tangent space(s) can be identified with the Lie algebra again. Hence both Jacobians are maps from the Lie algebra to itself. The signatures are
\verb|jacobian_exp(G, p, X, b)|
%\footnote{in-place variant \texttt{jacobian\_exp!($\mathfrak g$, J, p, X, b)}}
and \verb|jacobian_conjugate(G, g, X, b)|%
%\footnote{in-place variant \texttt{jacobian\_conjugate!($\mathfrak g$, J, g, X, b)}}
, where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.

@@

the main functions then are \verb|apply(a::GroupAction, g, p)|
%\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
to compute \(\sigma(g, p)\) for a group action \verb|a|, where $g\in \mathcal G$ and $p \in \mathcal M$, as well as its differential \verb|diff_apply(a::GroupAction, g, p, X)|
%\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
to compute \(D\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|
%\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \mathfrak g\).

@@

Note that while the group action $\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)$ “appends” a new action (w.r.t.~$g$) on the left, for the inverse we have
\begin{align*}
        \sigma^{-1}(g, \sigma^{-1}(h, p))
        &= \sigma(g^{-1}\cdot h^{-1}, p)
        = \sigma((h \cdot g)^{-1}, p)
        \\
        &= \sigma^{-1}(h \cdot g, p)
\end{align*}
for all \(g, h \in \mathcal{G}\). This is a so-called right action.
Similarly for a right group action the inverse is a left group action.
This is tanken into account when using \verb|apply|.

@@

If you want to construct nested Products of Lie groups, e.g.\ \((\mathcal G_1 \times \mathcal G_2) \times \mathcal G_3\), the binary operator would not work, since it automatically flattens the input; use \verb|ProductLieGroup(ProductLieGroup(G1, G2), G3)| can be used instead.

\subsubsection*{Power Lie groups}
Though mathematically the case of a product Lie Group, where $G=H$ is also (just) a product, computationally we can benefit from the special structure. Therefore the \verb|PowerManifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(G\) with itself.
As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(G\) with itself.

@@

Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast)\) and \(\mathcal{H} = (\mathcal N, \diamond)\) as well as a group action \(\sigma\colon \mathcal{G} \times \mathcal{H} \to \mathcal{H}\),
the \term{(left) semidirect product Lie Group} $\mathcal{G} \ltimes \mathcal{H} = (\mathcal{M} \times \mathcal{N}, \cdot)$ is defined via the group operation given by
\begin{equation*}
    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond \sigma(g_1, h_2))
\end{equation*}
This semidirect product Lie group is constructed calling \verb|LeftSemidirectProductLieGroup(G, H, a)|, where the \verb|GroupAction a| is optional, since one can define a default action by setting \verb|default_left_action(G, H)|.
With this default, also the shortcut \verb|G|\,$\ltimes$\,\verb|H| is available.
%
Similarly, the \term{right semidirect product Lie Group} $\mathcal{H} \rtimes \mathcal{G}$ is defined analogously with the positions of the groups swapped in the group operation and the default for the group action is given by \verb|default_right_action(G, H)|. This can be constructed via \verb|H|\,$\rtimes$\,\verb|G| then as well.

@@

The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing functionality.
For Lie groups there are currently two such decorators available

@@

The second representation is the tuple representation, where points are represented as tuples \(g' = (R, t)\) and the third is the tuple representation with the translation in the first component, i.e.\ \(g'' = (t, R)\). These two are available when \verb|RecursiveArrayTools.jl| is loaded and specifying the \verb|variant=:left| (default) or \verb|variant=:right|, respectively.
The first, matrix representation work on either \verb|variant|.

@@

>>>>

\begin{abstract}
    \href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}} is a
    Julia package that provides an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions. The package also offers a well-documented, performant, and well-tested library of such objects, with a focus on numerical computations in engineering.

@@

\subsection{Lie groups}
We denote a Lie group by \(\mathcal{G} = (\mathcal{M}, \cdot)\) where \(\mathcal{M}\) is a smooth manifold and \(\cdot\) is the group operation.
A smooth manifold \(\mathcal{M}\) is a topological space that is locally homeomorphic to an Euclidean space \(\mathbb{R}^n\) for some \(n \in \mathbb{N}\), but globally may have a different topology.
We call \(n\) the dimension of the manifold \(\mathcal{M}\), denoted by \(\dim(\mathcal{M}) = n\).
As an example, take the \(2\)-dimensional sphere
\begin{equation*}
\mathbb{S}^2 = \bigl\{ p \in \mathbb{R}^3 \bigm| \lVert p\rVert = 1\bigr\},
\end{equation*}
% MB: maybe it would be better to replace it with an example that can have a group structure? For example the 3-dimensional sphere?

which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
Finally we denote the tangent space at a point \(p \in \mathcal{M}\) by \(T_p\mathcal{M}\). This can be thought of as all “velocities” (direction and speed) in which a curve can “pass through” a point. Formally it is set the equivalence classes of derivatives of smooth curves. Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces
\begin{equation*}
    T\mathcal M = \dot\bigcup_{p \in \mathcal M} T_p\mathcal M
\end{equation*}
the \term{tangent bundle} of $\mathcal M$.
\\
A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) is a function that satisfies the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\). Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is, for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle. % chktex 9ok
The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$ when $\alpha \neq -\pi$ and $-\pi$ otherwise). %

\subsection{The Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} $\frak g$ of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} defined by $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(e)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
.
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constant vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ that consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.

@@

The interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
To generate a point in a specific representation by objects of type \verb|T|, one can use \verb|identity_element(G, T)|.
For example, a points on the special Euclidean group $\mathrm{SE}(n)$ can be represented either as a $(n+1) \times (n+1)$ matrix or a \verb|ArrayPartition| with rotation and translation stored separately.

@@

Alternatively to points, the same group structure can be defined over different topological manifolds. This is for example the case for the circle group \(\mathbb{S}^1\), which can be represented as angles, points on the unit circle in \(\mathbb{R}^2\) or as complex numbers with unit norm, all three of which are different manifolds to be used internally. To access the underlying manifold, the function \verb|base_manifold(G)| is provided.

The following functions are available and pass directly to the underlying manifold interface from \verb|ManifoldsBase.jl|:
\verb|is_approx(G, g, h)| to check for (approximate) equality of two points,
\verb|is_point(G, g; error=:none)| to check if a point is a valid point on the manifold, where the keyword can be used to throw an \verb|:error|, a \verb|:warn| or an \verb|:info|.
Similarly
\verb|manifold_dimension(G)| to get the dimension of the manifold,
\verb|project(G, q)|, to project a point onto the manifold, as well as %
%\footnote{in-place variant \texttt{project!(G, p, q)}.} 
\verb|rand(G)| and \verb|rand(rng, G)| %
%\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
to sample a random point from the manifold using a random number generator \verb|rng| also pass directly on to the manifold.
To access the underlying manifold of the Lie group one can use \verb|base_manifold(G)|.

@@

The identity element can be represented as a value of a special type, \verb|e = Identity(G)| or \verb|Identity(op)| where \verb|op| is the group operation of \verb|G|. This allows for dispatching on the type when defining functions and avoiding unnecessary allocations. If the actual value is needed, one can call
\verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group. The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|%
%\footnote{The in-place variant \texttt{identity\_element!(G, e::T)} infers the representation from the type \texttt{T} of \texttt{e}.}
.

The two central functions are \verb|compose(G, g, h)|
%\footnote{in-place variant \texttt{compose!(G, k, g, h)}.}
to compute the group operation \(g \cdot h\) for two points \(g, h \in \mathcal{G}\) and the inverse \verb|inv(G, g)|
%\footnote{in-place variant \texttt{inv!(G, h, g)}.}
which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).
Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|%
%\footnote{in-place variant \texttt{conjugate!(G, k, g, h)}.}
.

@@

Finally, there is a specific function for the differential of the conjugate at the identity $h=e$, called \verb|adjoint(G, g, X)|,
%\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.}
as well as the combinations \verb|inv_left_compose(G, g, h)|
%\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}}
for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|
%\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}
for computing $g \cdot h^{-1}$. All three fall back to the previously defined functions, but provide an interface to possibly implement more efficient variants in case such exist.

@@

Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code%
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}%
.
Keep in mind, that the Lie algebra is a vector space, so that addition, subtraction and scalar multiplication, as well as their broadcasted variants, are assumed to be defined in case you use an individual data type.

@@

Given a vector $c \in \mathbb R^{\dim(\mathfrak g)}$, we obtain the corresponding tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, B)| where \verb|B| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, B, T)| specifying the tangent vector type%
%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, B)} can infer the type \texttt{T} of \texttt{X} directly.}
.
Given a tangent vector $X \in \mathfrak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\mathfrak g$\verb|, X, B)|%
%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, B)}}
, where the basis is again optional.

@@

\subsubsection*{Push forward and pull back of tangent vectors}
To get a tangent vector $X_g \in T_g \mathcal G$ at point $g \in \mathcal G$ from its Lie algebra representation $X \in \mathfrak g$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$. This is implemented as \verb|push_forward_tangent(G, g, X)|%
%\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
.
Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\mathfrak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$. This is implemented as \verb|pull_back_tangent(G, g, Y)|%
%\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
.

\subsubsection*{Jacobians}
On a Euclidean space, the terms differential and Jacobian are often used interchangeably. Within differential geometry, the differential (or push forward) of a smooth map $f\colon \mathcal M \to \mathcal N$ between two manifolds is a map between the corresponding tangent spaces. We denote it by $Df(p)\colon T_p\mathcal M \to T_{f(p)}\mathcal N$, where $p \in \mathcal M$.
The differential is a linear map between the tangent spaces.
However, in the coordinate-free representation of tangent vectors, no matrix representation of this linear map is constructed.
But as soon as we choose bases for the tangent spaces and we represent tangent vectors in coordinates (cf.~\verb|get_coordinates| previously), we can represent the differential as a matrix, called the Jacobian matrix.

Currently for both \verb|exp| as well as \verb|conjugate|, the Jacobian is implemented. Note that for \verb|exp|, the domain is the Lie algebra. Since that is a vector space, its tangent space(s) can be identified with the Lie algebra again. Hence both Jacobians are maps from the Lie algebra to itself. The signatures are
\verb|jacobian_exp(G, p, X, B)|
%\footnote{in-place variant \texttt{jacobian\_exp!($\mathfrak g$, J, p, X, B)}}
and \verb|jacobian_conjugate(G, g, X, B)|%
%\footnote{in-place variant \texttt{jacobian\_conjugate!($\mathfrak g$, J, g, X, B)}}
, where \verb|B| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.

@@

The main functions then are \verb|apply(a::GroupAction, g, p)|
%\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
to compute \(\sigma(g, p)\) for a group action \verb|a|, where $g\in \mathcal G$ and $p \in \mathcal M$, as well as its differential \verb|diff_apply(a::GroupAction, g, p, X)|
%\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
to compute \(D_{\mathcal M}\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|
%\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
to compute \(D_{\mathcal G}\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \mathfrak g\).

@@

Note that while the group action $\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)$ “appends” a new action (w.r.t.~$g$) on the left, for the inverse we have
\begin{align*}
        \sigma^{-1}(g, \sigma^{-1}(h, p))
        &= \sigma(g^{-1}\cdot h^{-1}, p)
        = \sigma((h \cdot g)^{-1}, p)
        \\
        &= \sigma^{-1}(h \cdot g, p)
\end{align*}
for all \(g, h \in \mathcal{G}\). This is a so-called right action.
Similarly for a right group action the inverse is a left group action.
This is taken into account when using \verb|apply|.

@@

The binary operator $\times$ automatically flattens the input, so that the points are represented by non-nested \verb|ArrayPartition| objects.
If you want to construct products of Lie groups where points are represented by nested \verb|ArrayPartition| objects, for example \((\mathcal G_1 \times \mathcal G_2) \times \mathcal G_3\), use \verb|ProductLieGroup(ProductLieGroup(G1, G2), G3)| instead.

\subsubsection*{Power Lie groups}
Mathematically power groups are product Lie groups where all manifolds are the same. Computationally we can sometimes benefit by exploiting this special structure.
Therefore the \verb|PowerManifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(\mathcal G\) with itself.
As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(\mathcal G\) with itself.

@@

Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast)\) and \(\mathcal{H} = (\mathcal N, \diamond)\) as well as a group action \(\sigma\colon \mathcal{G} \times \mathcal{H} \to \mathcal{H}\),
the \term{(left) semidirect product Lie Group} $\mathcal{G} \ltimes \mathcal{H} = (\mathcal{M} \times \mathcal{N}, \cdot)$ is defined via the group operation given by
\begin{equation*}
    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond \sigma(g_1, h_2))
\end{equation*}
This semidirect product Lie group is constructed calling \verb|LeftSemidirectProductLieGroup(G, H, a)|, where the \verb|GroupAction a| is optional, since one can define a default action by setting \verb|default_left_action(G, H)|.
With this default, also the shortcut \verb|G|\,$\ltimes$\,\verb|H| is available.
%
Similarly, the \term{right semidirect product Lie Group} $\mathcal{H} \rtimes \mathcal{G}$ is defined analogously with the positions of the groups are swapped in the group operation and the default for the group action is given by \verb|default_right_action(G, H)|. This can be constructed via \verb|H|\,$\rtimes$\,\verb|G| then as well.

@@

The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing functionality.
For Lie groups there are currently two such decorators available: one for adding a custom metric structure (inner product) to the group and one for numerical validation.

@@

The second representation is the tuple representation, where points are represented as tuples \(g' = (R, t)\) and the third is the tuple representation with the translation in the first component, i.e.\ \(g'' = (t, R)\). These two are available when \verb|RecursiveArrayTools.jl| is loaded and specifying the \verb|variant=:left| (default) or \verb|variant=:right|, respectively.
The matrix representation works on either \verb|variant|.

@@

            align=center,
